Parameterization of the responses of subarctic European vegetation to key environmental variables for ozone risk assessment

### 2022-04-21 VegOzone.tex
Move reference to Appendix figures
Finish first part of section 3

To determine the temperature localization for coniferous trees, we used the time series of \chem{CO_2} uptake and temperature at observation sites in southern ($\rightarrow$ \emph{cold}) and northern ($\rightarrow$ \emph{subarctic}) Finland by \citet{TB:Kolari2007}. From these, we estimated an optimal temperature interval ($10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$). The same interval was assumed for deciduous trees.

@@

Soil texture influences water availability and thus $\mathrm{POD_y}$ calculations in $\mathrm{DO_3SE}$ \citep{ACP:Bueker2012}. The soil at Svanhovd is characterized in an ICP~Forest plot survey as gleyic cambisols according to the Food and Agriculture Organization of the United Nations (FAO) soil classification system and as eluviated dystric brunisol in the ICP~Forest soil classification system (V. Timmerman, personal communication, June 2020). The clay:sand:silt content is estimated to range from 12\,\%:61\,\%:27\,\% at the top layer to 16\,\%:59\,\%:25\,\% in about $1\,\unit{m}$ depth (\href{https://soilgrids.org/}{soilgrids.org}, last accessed April 2022 for location $30.0334\,\unit{^\circ E}$, $69.4549\,\unit{^\circ N}$) Within the $\mathrm{DO_3SE}$ soil parameterization, the sandy loam texture describes the upper $60\,\unit{cm}$, where the plant roots are found, of this soil type best.

@@

All bespoke temperature parameters are listed in Table~\ref{tab:sensitivity_tests_temp}.

(for deciduous and coniferous trees refer Appendix~Figs.~\ref{fig:f_temp_spruce}b and \ref{fig:f_temp_birch}b)

All derived parameters are tabulated in Table~\ref{tab:sensitivity_tests_light}.

We presume that perennial vegetation will likely maximize carbon acquisition during the short subarctic growing season. Because the MM version of the $\mathrm{DO_3SE}$ model currently does not simulate net photosynthesis ($A_\mathrm{n}$), we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune the temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ to acclimate the growth potential to the local climate. For details regarding these functions see Section~\ref{subsec:model_describtion}. Further, we presume that $f_\mathrm{VPD}$ and $f_\mathrm{SWP}$ suit our vegetation types and no acclimation is necessary for these. But acknowledge that these are very important factors as shown by~\citet{ACP:Bueker2012}.

As indicated above, a better acclimation to prevailing climate conditions is characterized by a higher average $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$. To compare our bespoke parameterizations, we propose the following metric. We calculate an average relative stomatal conductance $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal}) with $f_\mathrm{SWP}=1$) at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) for the whole record of key environmental variables' data and apply mean and standard deviation. We presume that $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ around noon (highest light intensity) is a good proxy for \chem{CO_2} uptake efficiency (growth). In addition, a small deviation between noon and morning and a low standard deviation would indicate higher robustness to variability in growing conditions. Based on these criteria, we identify \emph{subarctic}--PPFD0.8 as the best parameterization for all PFTs (Fig.~\ref{fig:javis_func_opt_mean}). As expected due to the small adjustments, coniferous trees display the smallest differences between the different parameterizations, while the differences for perennial grassland are substantial as a response to the proposed temperature acclimation.

@@

Start ($A_\text{start}$) and end ($A_\text{end}$) of the GS for each PFT are usually given in \unit{doy} (Table~\ref{tab:sensitivity_tests_gs})

>>>>

To determine the temperature localization for coniferous trees, we used the time series of \chem{CO_2} uptake and temperature at observation sites in southern ($\rightarrow$ \emph{cold}) and northern ($\rightarrow$ \emph{subarctic}) Finland by \citet{TB:Kolari2007}. From these, we estimated an optimal temperature interval ($10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$). The same interval was assumed for deciduous trees. For deciduous and coniferous trees refer to Appendix~Figs.~\ref{fig:f_temp_spruce}b and \ref{fig:f_temp_birch}b), respectively.

@@

Soil texture influences water availability and thus $\mathrm{POD_y}$ calculations in $\mathrm{DO_3SE}$ \citep{ACP:Bueker2012}. The soil at Svanhovd is characterized in an ICP~Forest plot survey as gleyic cambisols according to the Food and Agriculture Organization of the United Nations (FAO) soil classification system and as eluviated dystric brunisol in the ICP~Forest soil classification system (V. Timmerman, personal communication, June 2020). The clay:sand:silt content is estimated to range from 12\,\%:61\,\%:27\,\% at the top layer to 16\,\%:59\,\%:25\,\% in about $1\,\unit{m}$ depth (\href{https://soilgrids.org/}{soilgrids.org}, last accessed April 2022 for location $30.0334\,\unit{^\circ E}$, $69.4549\,\unit{^\circ N}$) Within the $\mathrm{DO_3SE}$ soil parameterization, the sandy loam texture describes the upper $60\,\unit{cm}$, where the plant roots are found, of this soil type best. We presume that no localization of the water vapor pressure deficite $f_\mathrm{VPD}$ and soil water potential $f_\mathrm{SWP}$ response functions is necessary for our species. But acknowledge that these are very important factors in the modeling of ozone dose as pointed out by~\citet{ACP:Bueker2012}.

@@

We presume that perennial vegetation will likely maximize carbon acquisition during the short subarctic growing season. Because the MM version of the $\mathrm{DO_3SE}$ model currently does not simulate net photosynthesis $A_\mathrm{n}$, we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011}. A higher average $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ ratio, thus, indicates a higher growth potential in the local climate.

To compare our localized parameterizations, we propose the following metric. From the long term of meteorological data, we select to distinct periods noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and morning ($5-9\,\unit{am}$ local time). For these, we calculate the average relative stomatal conductance $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal}) and assume $f_\mathrm{SWP}=1$). We presume that $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ around noon (the highest light intensity) is a good proxy for \chem{CO_2} uptake efficiency (growth). In addition, a small deviation between noon and morning and a low standard deviation indicates higher robustness of this localization to variability in growing conditions. From left to right in Fig.~\ref{fig:javis_func_opt_mean}, results using the subarctic, cold, and MM $f_\mathrm{T}$ are shown marked by the alternating background color. Within each of these groups $f_\mathrm{light}$ categories are indicated by PPFD0.8 and PPFD1.2, respectively. The standard deviation is denoted with error bars.

Based on these criteria, we identify \emph{subarctic}--PPFD0.8 as the parameterization with the highest score for all PFTs. The differences between the localized parameterizations and the MM are smallest for coniferous trees. This indicates that the MM Boreal parameterization is well suited also for subarctic climate at Svanhovd. For perennial grassland, the difference compared to MM is largest suggesting that the MM Central European parameterization is not appropriate in a subarctic climate.

@@
